---
title: "explore"
author: "Hong Jiang"
date: "October 22, 2015"
output: html_document
---

SW: you have the right pieces, but they are not well organized. your output is confusing, you've dropped labels in a lot of cases, and you repeat the summary measures for the numerical and categorical variables. you need to look at the absolute value of the correlation coefficient, not just positive values. you also should explicitly call `return`. separate note, good job dealing with binary variables.

```{r}
require(grid) #load grid package for plots
require(ggplot2) #load ggplot2 package for plots

explore<-function(dataframe,binsize,cor_threshold)
  { #Define the explore function with input data dataframe, binsize and cor_threshold
  
  #Question 1: we first extract the numeric varibales and then use loop for histgram by ggplot()
  nums<-dataframe[sapply(dataframe,is.numeric)] #using sapply() to find all numeric column and put them into variable nums
  histlist<-list() #create a varable histlist to put histograms which is empty now
  for (i in 1:length(binsize)){ #go through from the first number to the last in binsize
    for (j in 1:(ncol(nums))){ #go through from the first column to the last in nums
      binw<-(max(nums[,j])-min(nums[,j]))/binsize[i] #calculating binwidth for histogram 
      histlist<-ggplot(nums,aes(x=nums[,j]),environment=environment()) #using ggplot to draw the plot for every numerical variable,use environment parameter
      histlist<-histlist+geom_histogram(colour="blue",fill="blue",binwidth=binw)+labs(x=colnames(nums)[j])+geom_vline(xintercept=mean(nums[,j]),colour="red") #add blue histogram,label x, then draw a vertical red line at the mean
      print(histlist) #output histograms 
      
      print(histlist+aes(y=..density..)+labs(y="density")+geom_density()) #output histograms using density and label the y label 
    }


  
  
  #Question 2: 
  factors<-dataframe[sapply(dataframe,is.factor)] #using sapply() to find all factor columns 
  binarys<-data.frame(matrix(ncol=0, nrow=nrow(dataframe))) #create a data frame for binarys 
  a=1 #create a variable a and use it to write binarys columns into data frame
  for (i in 1:ncol(dataframe)){ #loop
    if (sum(dataframe[,i]==1)+sum(dataframe[,i]==0)==nrow(dataframe)){ #use if() and sum() to check if there are columns only have 0s and 1s
      binarys<-data.frame(binarys,dataframe[,i]) #store the binary variables into binarys variable
      names(binarys)[a]=colnames(dataframe)[i] #make sure the name of the column won't change in new data frame
      a=a+1 #add 1 to a in order to go to the next column in binarys
    }#finish if
  } #finish for loop
  fnb<-data.frame(factors,binarys) #create a data frame fnl and put factors and logicals in it
  plotlist<-list() #create a variable plotlist for a list of plots, make it empty, we will use it to put all plots
  for(i in 1:ncol(fnb)){ #using for loop to go through all variables in fnl
    plotlist[[i]]<-ggplot(fnb,aes_string(x=colnames(fnb)[i]),environment=environment())+geom_bar(colour="gray",fill="grey")+ggtitle(paste(colnames(fnb)[i],"distribution")) #put a gray bar graph for ith column in fnl into plotlist[[i]], label xlabel and write title, and use use environment parameter
  } #finish for loop
  print(plotlist)#output plotlist
  
  
  
  #Question 3:
  Pair_of_variables<-c() #create a variable Pair_of_variables
  rsqr<-c() #create a variable rsqr, will store r-square value in it
  n=1 #create a variable n that will represent position in Pair_of_variables and rsquared and make it equal to 1 first
  for (i in 1:(ncol(nums)-1)){ #use for loop
    for (j in (i+1):ncol(nums)){ #use for loop to go through from the i+1th column name to the last in dataframenum
      Pair_of_variables[n]<-paste(colnames(nums)[i],"-",colnames(nums)[j],sep="") #using paste() to write pair of variable names in a single string separated by a -
      rsqr[n]<-summary(lm(nums[,i] ~ nums[,j]))$r.squared #using summary() and lm() to get the r-square value between two varaibles
      n=n+1 #add 1 to n in order to go to the next positions in Pair_of_variables and rsq
    } #finish second for loop
  }#finish first for loop
  
  #Question 4bii
  newdata<-data.frame(Pair_of_variables, rsqr) #create a data frame newdata that has Pair_of_variables and rsqr 
 
   print(newdata) #output newdata 
  #Question 4a:
  tablelist<-list() #create a variable tablelist for a list of tables
  for (i in 1:ncol(fnb)){ #using for loop to go through all categoricals and binary variables in dataframe
    tablelist[[i]]<-as.data.frame(table((fnb)[,i])) #use table() to give the counts of ith column in fnb, convert it to a data frame and put into tablelist[[i]]
    names(tablelist[[i]])[1]=colnames(fnb[i]) #use names() to get the variable name in the corresponding column name
  } #finish for loop
  print(tablelist) #output 
  #Question 4bi: 
  thetable<-list() #create a variable  for a list of table
  for (i in 1:ncol(nums)){ #for each numeric column in the data frame
  thetable[[i]] <- summary(nums[,i]) 
  # SW: when you print this, you lose the variable names, so it is not helpful
  #let the summary table of ith column to be the ith element in thetable
  }#finish foor loop
  print(thetable) #output 
  #Question 4biii:
  Pairofvariables<-c() #create a variable Pairofvariables
  Pearson_cor_coeff<-c() #create a variable Pearson_cor_coeff
  n=1 #create a variable n that will represent position in Pairofvariables and Pearson_cor_coeff
  for (i in 1:(ncol(nums)-1)){ #use loop to go through column name 
    for (j in (i+1):ncol(nums)){ #use loop to go through from the i+1th column name to the last in nums
      if(cor(nums[ ,i],nums[ ,j],method="pearson")>cor_threshold){ # check if the pearson correlation coefficient between ith column and jth column in nums larger than the input value cor_threshold
      Pairofvariables[n]<-paste(colnames(nums)[i],"-",colnames(nums)[j],sep="") #paste() to write pair of variable names in a single string separated by a -, and put into Pair_of_variables
        Pearson_cor_coeff[n]  <- cor(nums[ ,i],nums[ ,j],method="pearson") #cor() to calculate the Pearson correlation coefficient and store the result in Pearson_cor_coeff
        n=n+1 #add 1 to n in order to go to the next positions 
      }#finish if
    } #finish second for loop
  }#finish first for loop
  perdata<-data.frame(Pairofvariables, Pearson_cor_coeff) #create a data frame called including Pairofvariables and Pearson_cor_coeff 
  print(perdata) #output perdata
  }
}
```

```{r}
#question 5 test:
data(diamonds) #load data
logicalcol<-c() #create variable logicalcol
ratioT=length(mtcars$vs[mtcars$vs==1])/length(mtcars$vs) #Calculate the ratio of 1 in the mtcars$vs variable, and put it in variable ratioT
trail<-rbinom(nrow(diamonds),1,ratioT) #use rbinom() to randomly input 0 and 1 into trail based on ratioT, and let the length of trail equal to the length of diamonds
for (i in 1:(length(trail))){ #oop to go through  trail
  if (trail[i]==1){ #if() to check if trail[i] equals to 1
      logicalcol[i]=TRUE #if trail[i] equals to 1, write TURE in logicalcol[i]
      } #finish if(){}
  else{
    logicalcol[i]=FALSE #if trail[i] doesn't equal to 1, write FALSE in logicalcol[i]
    } #finish eles{}
} #finish for loop
newdiamonds<-data.frame(diamonds,logicalcol) #create a new data frame called newdiamonds and put diamonds and logicalcol into it
# SW: what is this column? give it a descriptive name.

explore(newdiamonds, c(5,20,50), 0.25) #test explore() using newdiamonds
explore(mtcars,c(5,20,50),0.25) #test explore() using mtcars
```
